### 2022-10-18 response2.tex
continued doing referee revisions

\documentclass[12pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=20mm]{geometry}
\usepackage{amsmath,color,amsthm,enumerate,amsfonts,mathtools}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{paralist}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{todonotes}

\newcommand{\done}{\textcolor{blue}{Done.}}
\newcommand{\changed}{\textcolor{blue}{Revised accordingly.}}
\newcommand{\tba}{\textcolor{red}{To be addresssed.}}

\newenvironment{response}{\color{blue}}{}

\setlength{\parindent}{0ex}
\setlength{\parskip}{1ex}

\begin{document}

\subsection*{Graph Product Structure for Non-Minor-Closed Classes\\
Vida Dujmovi\'c, Pat Morin, David R. Wood\\
Journal of Combinatorial Theory Series B (JCTB9758)\\
Response to Referee Report received 4th October 2022}

Thanks to the referee for thoroughly checking the paper, and for highlighting several places for improvement. Here we respond to the referee's comments.

\hrulefill

1. In the proof of Theorem 7, I see no reason why we should have $a = x_r$
(or $x'=x_0$, for that matter). I think you need to use the fact that $v_r$ is
in $S_a$. (Also, if we did have that $a = x_r$, then $j$ would be $r$, so maybe
there is a different mistake, or the intention is different.) It might be
helpful to define $H^+$ first and then have a separate claim which shows
how to use paths in $H$ to get paths in $H^+$.

\tba

2. Are the nations incident to a vertex counted with multiplicity in dmap graphs? The definition does not say so, but the proof of Lemma 27 (as currently stated) assumes not only this, but that the embedded graph under consideration is 2-connected. It also seems that frames are loopless, so slightly more care is called for when they are constructed.
It may be worth noting explicitly that you allow d-framed graphs to
have edges outside the outer face of their frame - I was unsure about
this given the phrase “interior of the face”.

\tba

3. Sperner’s Lemma is usually stated for simple graphs (as is the case
in the source given) so there should be a reduction argument. It is
more than just removing parallel edges because there may be separating
cycles of length two.

\tba

Other comments:

1. There are some proofreading things to do - some lines overflow, there
are citations on a new line (instead of using tilde), and there are typos
like missing words (page 3, line 7, should read “H is a planar graph”).

\begin{response}
	We have fixed ``$H$ is a planar graph'', but do not see the point in fixing minor line overflows, since the paper will be reformatted in the journal style anyway.
\end{response}

2. Page 4, second bullet point. It says that you derive product structure
theorems for string graphs, which is clearly not possible. Please use
more precise language. This comment was in the previous review as
well. Some but not all instances have been corrected.

\begin{response}
We do not understand this comment. Why can't there be a product structure
theorem for string graphs?
\end{response}

3. page 6, line -6: The elements of P are not ordered so it is not a sequence.

\tba

4. Page 7, line 6: It is confusing to say that Sx “is defined by vertices
that participate in x” because it suggests that Sx is the set of vertices
which participate in x. Moreover, for the proven upper bound to be
enough, you need to show that the set of all nodes which participate in
x induces a subtree of T (and thus that a(v) is in this set).

\tba

5. The paper regularly says things like “let P be an H-partition” without
saying what H is. Typically this is understandable, but it could be
written more clearly. For instance, on page 8, line 2, it says “let Yxi
be the part in Y...” without saying what xi
is (and similarly in other
places). This shouldn’t be a serious hang-up for anyone, but still, it
would be better to acknowledge technical inaccuracies.

\tba

6. page 8, Claim 3. The definition of Cx, a subset of V (J), includes things
which may not be vertices of J. This, of course, is not a real problem,
but the proof treats the fact that V (J) may be a proper subset of
V (T) sloppily. Also, the notation T[x] is unfortunately ambiguous - it
is strange to need to define something which has already been defined.

\tba

7. page 16, line 6: Say why that is a convenient way to establish the
planarity of H. (It does not actually say here that any such graph G
is planar.)

\begin{response}
	Added the desired sentence.
\end{response}

8. page 16, line 8. I would not call these “BFS spanning forests rooted
at V0” since they need not be BFS spanning forests (unless the same
terminology has already appeared elsewhere?).

\tba

9. page 16, Lemma 24. The lemma is rather long to state and I would
imagine that anyone reading it has already read the proof of the product
structure theorem for planar graphs. So I would add a sentence or two
beforehand about how the set-up relates to that proof/tripods.

\tba

10. page 16, Lemma 24: Add “non-empty” before “sets $P_1,\dots,P_k$”.

\done

11. page 16, line -2. “$P_1, P_2, P_3$” should be $P_1,\dots,P_k$.

\done

12. page 17, top. In the first two cases there may not be a vertex v3 defined
as we could have p = 2.

\tba

13. page 17, right after 3. The end of Qi should be the same as the vertex w
chosen in case 3 because there may be multiple “appropriate” vertices.
It should also really say what “appropriate” means.

\tba

14. page 17, line -9. “$\{x_1, x_2, x_3\}$” should be “$\{v_1, v_2, v_3\}$”

\tba

15. page 19, second to last paragraph. Add a sentence about why such
a forest exists. The definition of a BFS layering is missing. A reader
may expect that you will just use the assumptions about H and P
that are stated, so it would be better to say something like “This gives
an H-partition... and conditions (i), (ii), and (iii) of Lemma 24 are
satisfied”.

\tba

16. page 20, first paragraph of Lemma 25. Well, parallel edges can appear
on the boundary of a common face if they were there before. Add a
sentence about why the process ends.

\tba

17. page 20, second to last paragraph. Figure 4 is mentioned several paragraphs before the graph which it depicts is defined. The graph $G_0$ in
this paragraph should be $G'$, and check for this throughout.

\tba

18. page 21, Lemma 27. Should be “labelled a nation or a lake” in the
first paragraph. In the third paragraph, the graph $\hat{G}_1$ should be “edge-maximal” or similar.

\tba
\end{document}